---
title: "linear model from Web"
author: "Taylor Stohl"
date: " Summer 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot


```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```


```{r}
basicNN + geom_point()
```



  
The y-axis is SIMS and the x-axis is ARM. This plot shows that there is a positive linear correlation. Theres's a positive correlation between ARM and SIMS displayed in this plot above.







### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The value is 0.68, which is a positive number that indicates the line will travel upwards towards the right.






### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

This model is better than thr one for GRIP. The Adjusted R-Square is bigger, which means more errors were reduced.


If you go 1.226 up or down that is the prediction interval.

 The residual standard error represents the standard deviation of the  error points
In this a linear regression model. the numbers that should be considered to create the prediction equation are the estimate coefficients. To predict SITS based on ARM strength, the intercept, -4.1 would be added to .05,ARM and then multiplied by ARM. 
This equation for the prediction of SIMS would look like this:-4.1 + .05 * ARM.
The formula needed is y = mx + b to create a line. The bigger the Adjusted R-squared, the better the model because it means that many of the errors were fixed. The degrees of freedom will be taken away from the Multiple-R squared and then what will be left is the Adjusted R-squared.
In the adjusted R Square, the error was reduced by 46%(0.467).
The adjusted R-squared represents is the reduction of the error from the mean model on a percentage basis.

The 145 degrees of freedom is how many data points are displayed in the plot. 
The line is to be estimate where it hits the y-axis. Each data will change the degree of freedom. Every time you add more stuff to the model, the Multiple R-squared will go down a small percentage.







#### Predict at target point





#### scatterplot with model fit
  

## Model 2 SIM~GRIP

### Now add in scatterplot


### Numerical results 


### Inferential  (Build model.2)
  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

